Generality in food webs scales with species richness but not with latitude

### 2015-06-24 manuscript/coverletter.tex
Updated coverletter, results in main text

\begin{letter}{\bf RE: \mytitle\\
               ~~~~~~~Cirtwill, Alyssa R; Stouffer, Daniel B; Romanuk, Tamara N\\
               \vspace*{20pt}
               Professor Spencer C.H. Barrett\\
               Editor-in-Chief, Proceedings of the Royal Society B\\
               6-9 Carlton House Terrace\\
               London, UK\\
               SW1Y 5AG\\
                }

@@

Our study aims to combine two well-established but disparate ecological themes, namely the analysis of macroecological gradients and the search for general laws governing the structure of communities, as measured through food webs (networks of who eats whom). In particular, we were interested in the possibility of a link between the well-known latitudinal gradients in species richness of some taxa (e.g., birds, land plants) and the scaling relationships between species richness and food-web structure. With regard to the former, one  hypothesized mechanism for greater species richness in the tropics is the latitude-niche-breadth hypothesis, which predicts that the greater stability of temperature and productivity in the tropics will permit species to evolve narrower niches and therefore more finely partition habitats. If this hypothesis holds true, then tropical species should be more specialized than their high-latitude counterparts.

In terms of a food web, this means that we would naively expect that tropical species will have relatively fewer feeding links than polar species. However, the number of feeding links in which a species is involved is known to scale with the number of species in the community food web following a power law. Therefore we also expect that species in tropical food webs will participate in many feeding links by virtue of their speciose communities. This apparent paradox can be resolved by adding an effect of latitude to the exponent of the power law. We tested for the impact of such an exponent, as well as potential interactions between the effects of latitude and ecosystem type on scaling laws.

We found that, in general, there was little effect of latitude on scaling relationships between diet breadth and species richness. If diet breadth may be taken as a measure of a species' Eltonian (biotic) niche, this means that our results are inconsistent with the latitude-niche-breadth hypothesis. Ecosystem type, on the other hand, had a greater impact on scaling exponents, suggesting that different processes may structure food webs in different habitats.

Our results are the pinnacle of interestingness and you should publish them before we decide to give them to Nature instead. Everyone will love us and then the world will shower me with options for my future. That's how this works, right?

We believe that our study will be of interest not only to theory-focused biologists but also to general readers
of \emph{\myjournal} as it deals with global trends in the structure of ecological communities. 
Specifically, we address the impact of latitude and habitat type on known scaling relationships between empirical
food-web structure and species richness. Although latitude (as a proxy for temperature, seasonality, and primary 
productivity) is expected to drive trends in niche size (here measured as the number of feeding links for any 
given species), we instead show that habitat type has the stronger influence on scaling between niche size and 
species richness. This suggests that there may indeed be general rules governing the structure of food webs but 
that these rules differ between different habitat types.

Our paper thus contributes to the discussions of both the latitude-niche-breadth hypothesis and scaling
laws in food-web structure, as well as being potentially applicable to researchers exploring community 
structure and specialization in a variety of different systems. We hope you will agree that this broad relevance
and the novelty of our results justify our manuscript's inclusion in \emph{\myjournal}.

>>>>

\begin{letter}{\bf Professor Spencer C.H. Barrett\\
               Editor-in-Chief\\
               Proceedings of the Royal Society B\\
               6-9 Carlton House Terrace\\
               London, UK\\
               SW1Y 5AG\\
                }

@@

Our study aims to combine two well-established but disparate ecological
themes, namely the analysis of macroecological gradients and the search for
general laws governing the structure of food webs (networks of who eats whom
within a community). In particular, we were interested in the possibility of a
link between the well-known latitudinal gradients in species richness of some
taxa (e.g., birds, land plants) and the scaling relationships between species
richness and food-web structure. With regard to the former, one  hypothesized
mechanism for greater species richness in the tropics is the latitude-niche-
breadth hypothesis, which predicts that the greater stability of temperature
and productivity in the tropics will permit species to evolve narrower niches
and therefore more finely partition habitats. If this hypothesis holds true,
then tropical species should be more specialized than their high-latitude
counterparts. In terms of a food web, this means that we would naively expect
that tropical species will have relatively fewer feeding links than polar
species. However, the number of feeding links in which a species is involved
is known to scale with the number of species in the community food web
following a power law. Therefore we also naively expect that species in
tropical food webs will participate in many feeding links by virtue of their
speciose communities. This apparent paradox can be resolved by adding an
effect of latitude to the exponent of the power law. We tested for the impact
of such an exponent, as well as potential interactions between the effects of
latitude and ecosystem type on scaling laws.

In general we found little evidence of an effect of latitude on scaling
relationships between diet breadth and species richness, implying that
latitudinal variation in food-web structure may be driven purely by
latitudinal gradients in species richness. If diet breadth may be taken as a
measure of a species' Eltonian (biotic) niche, this means that our results are
inconsistent with the latitude-niche-breadth hypothesis. Ecosystem type, on
the other hand, had a greater impact on scaling exponents, suggesting that
different processes may structure food webs in different habitats. Indeed,
freshwater food webs departed from the overall trend and showed strong effects
of latitude on scaling relationships. In these food webs alone was the
latitude-niche-breadth hypothesis supported by our results.

By linking macroecology and food-web ecology, we believe that our study will
be of interest to ecologists, theory-focused biologists, and to general
readers of \emph{\myjournal}. Our results are potentially applicable to
researchers exploring community structure and specialization in a variety of
different systems and also provide a new perspective on the debate over the
latitude-niche-breadth hypothesis. We hope you will agree that this broad
relevance and the novelty of our results justify consideration of our
manuscript's inclusion in \emph{\myjournal}.

### 2015-06-24 manuscript/latitudepaper.tex
Updated coverletter, results in main text

  Link density, generality, and vulnerability each increased with increasing
  species richness ($\beta_0$=0.643, $p$\textless0.001; $\beta_0$=0.599,
  $p$\textless0.001; and $\beta_0$=0.643, $p$\textless0.001, respectively;
  Fig~\ref{props_v_lat}).For estuarine, marine, and terrestrial food webs the
  strength of this scaling did not vary with latitude
  ($\beta_{Latitude}$=-0.001, $p$=0.232 for link density;
  $\beta_{Latitude}$=-0.001, $p$=0.257 for generality; and
  $\beta_{Latitude}$=-0.001, $p$=0.233 for vulnerability; Fig.~\ref{S}). In
  lake food webs, however, the scaling of each property was stronger towards
  the poles ($\beta_{Latitude:Lake}$=0.004, $p$=0.006;
  $\beta_{Latitude:Lake}$=0.006, $p$\textless0.001; and
  $\beta_{Latitude:Lake}$=0.005, $p$=0.005, respectively). In stream food
  webs, generality increased more rapidly towards the poles
  ($\beta_{Latitude:Stream}$=0.007, $p$\textless0.001) while link density and
  vulnerability did not vary with latitude.

  Unlike species richness, only generality showed an overall increase with increasing
  proportions of basal resources in a web ($\beta_0$=-0.024, $p$=0.838;
  $\beta_0$=0.333, $p$=0.012; and $\beta_0$=-0.024, $p$=0.839 for link
  density, generality, and vulnerability, respectively;
  Fig.~\ref{props_v_lat}). In each case, however, there were substantial
  effects of latitude on the scaling relationships. Thus, the scaling
  exponents for link density and vulnerability were negative for estuarine,
  lake, and stream communities and became more negative towards the poles
  ($\beta_{Latitude}$=-0.007, $p$=0.005 and $\beta_{Latitude}$=-0.007,
  $p$=0.005, respectively; Fig.~\ref{BIT}). For the scaling of link density and
  vulnerability in marine and terrestrial communities, the overall effect of
  latitude on scaling was positive ($\beta_{Latitude:Marine}$=0.009,
  $p$=0.037; $\beta_{Latitude:Terrestrial}$=0.011, $p$=0.005 and
  $\beta_{Latitude:Marine}$=0.009, $p$=0.036;
  $\beta_{Latitude:Terrestrial}$=0.011, $p$=0.005, respectively) such that
  there was little overall scaling relationship. In contrast, the scaling
  exponent of generality with the proportion of basal resources was positive
  near the equator and negative near the poles in estuarine, lake, and stream
  communities ($\beta_{Latitude}$=-0.009, $p$=0.002) while the scaling
  exponents for marine and terrestrial communities were not significantly
  different from zero ($\beta_{Latitude:Marine}$=0.011, $p$=0.021;
  $\beta_{Latitude:Terrestrial}$=0.014, $p$=0.003).

  As with species richness, link density, generality, and vulnerability
  generally increased with the proportion of intermediate consumers in a web
  ($\beta_0$=1.74, $p$\textless0.001; $\beta_0$=1.92, $p$\textless0.001; and
  $\beta_0$=1.75, $p$\textless0.001 respectively; Fig.~\ref{props_v_lat}). The
  strength of this scaling decreased towards the poles in estuarine, lake,
  marine, and stream communities in each case ($\beta_{Latitude}$=-0.012,
  $p$=0.007; $\beta_{Latitude}$=-0.018, $p$\textless0.001; and
  $\beta_{Latitude}$=-0.012, $p$=0.007, respectively; Fig~\ref{BIT}). In
  terrestrial communities, however, scaling was always stronger towards
  the poles ($\beta_{Latitude:Terrestrial}$=0.016, $p$=0.023 for link density;
  $\beta_{Latitude:Terrestrial}$=0.022, $p$=0.004 for generality; and
  $\beta_{Latitude:Terrestrial}$=0.016, $p$=0.023 for vulnerability,
  respectively).

  In contrast, link density, generality, and vulnerability decreased as the
  proportion of top predators increased ($\beta_0$=-0.511, $p$\textless0.001;
  $\beta_0$=-0.418, $p$\textless0.001; and $\beta_0$=-0.511, $p$\textless0.001
  respectively; Fig~\ref{props_v_lat}). For estuarine, lake, and marine
  food webs the strength of this scaling did not
  vary significantly with latitude ($\beta_{Latitude}$=-0.001, $p$=0.572 for
  link density; $\beta_{Latitude}$=-0.001, $p$=0.566 for vulnerability; the
  best-fit model for generality did not include any terms for latitude in
  any ecosystem type). In stream and terrestrial food webs, scaling was more
  strongly negative in the tropics and did not differ from zero near the poles,
  although this trend was only significant in streams
  ($\beta_{Latitude:Stream}$=0.011, $p$=0.002 and
  $\beta_{Latitude:Terrestrial}$=0.005, $p$=0.136 for link density;
  $\beta_{Latitude:Stream}$=0.012, $p$=0.002 and
  $\beta_{Latitude:Terrestrial}$=0.005, $p$=0.133 for vulnerability; Fig~\ref{BIT}).

@@

  The lack of a latitudinal gradient in species richness in this dataset contrasts
  strongly with other studies~\citep{Schemske2009,Macpherson2002,Kaufman1995}. As numbers of species and links included in a food web
  vary strongly with sampling effort as well as with the underlying diversity of the study
  area, it is possible that the lack of latitudinal trends here is a result of researchers
  tending to expend similar amounts of sampling effort across studies. This could result in
  food webs describing species-rich tropical communities omitting more species and links
  than species-poor arctic communities. In addition, it is worth noting that gradients in
  species richness are generally measured for a single taxonomic group at a time (e.g.~\citep{Kaufman1995}).
  It is possible that these taxa are not well-represented in our food webs and that the
  dominant taxa in them do not have an underlying latitudinal gradient in richness. 
  In either case, the lack of a strong association between
  species richness and latitude in any ecosystem type
  means that any effect of latitude on other scaling 
  relationships is not being driven by an underlying 
  latitudinal gradient in species richness.
  The lack of confounding effects of latitude allows us to 
  more clearly assess effects of latitude on scaling with 
  species richness and proportions of species in different 
  trophic levels.

>>>>

  Link density, generality, and vulnerability each increased with increasing
  species richness ($\beta_0$=0.644, $p$\textless0.001; $\beta_0$=0.557,
  $p$\textless0.001; and $\beta_0$=0.644, $p$\textless0.001, respectively;
  Fig~\ref{props_v_lat}).For estuarine, marine, and terrestrial food webs the
  strength of this scaling did not vary with latitude
  ($\beta_{Latitude}$=-0.001, $p$=0.206 for link density;
  $\beta_{Latitude}$=-0.001, $p$=0.373 for generality; and
  $\beta_{Latitude}$=-0.001, $p$=0.207 for vulnerability; Fig.~\ref{S}). In
  lake food webs, however, the scaling of each property was stronger towards
  the poles ($\beta_{Latitude:Lake}$=0.005, $p$=0.005;
  $\beta_{Latitude:Lake}$=0.006, $p$\textless0.001; and
  $\beta_{Latitude:Lake}$=0.005, $p$=0.005, respectively). In stream food
  webs, generality increased more rapidly towards the poles
  ($\beta_{Latitude:Stream}$=0.007, $p$\textless0.001) while link density and
  vulnerability did not vary with latitude.

  Unlike species richness, only generality showed an overall increase with increasing
  proportions of basal resources in a web ($\beta_0$=-0.030, $p$=0.801;
  $\beta_0$=0.328, $p$=0.013; and $\beta_0$=-0.030, $p$=0.802 for link
  density, generality, and vulnerability, respectively;
  Fig.~\ref{props_v_lat}). In each case, however, there were substantial
  effects of latitude on the scaling relationships. Thus, the scaling
  exponents for link density and vulnerability were negative for estuarine,
  lake, and stream communities and became more negative towards the poles
  ($\beta_{Latitude}$=-0.007, $p$=0.006 and $\beta_{Latitude}$=-0.007,
  $p$=0.005, respectively; Fig.~\ref{BIT}). For the scaling of link density and
  vulnerability in marine and terrestrial communities, the overall effect of
  latitude on scaling was positive ($\beta_{Latitude:Marine}$=0.009,
  $p$=0.038; $\beta_{Latitude:Terrestrial}$=0.012, $p$=0.003 and
  $\beta_{Latitude:Marine}$=0.009, $p$=0.037;
  $\beta_{Latitude:Terrestrial}$=0.012, $p$=0.003, respectively) such that
  there was little overall scaling relationship. In contrast, the scaling
  exponent of generality with the proportion of basal resources was positive
  near the equator and negative near the poles in estuarine, lake, and stream
  communities ($\beta_{Latitude}$=-0.009, $p$=0.002) while the scaling
  exponents for marine and terrestrial communities were not significantly
  different from zero ($\beta_{Latitude:Marine}$=0.011, $p$=0.021;
  $\beta_{Latitude:Terrestrial}$=0.014, $p$=0.002.

  As with species richness, link density, generality, and vulnerability
  generally increased with the proportion of intermediate consumers in a web
  ($\beta_0$=2.10, $p$\textless0.001; $\beta_0$=0.592, $p$=0.006; and
  $\beta_0$=0.843, $p$\textless0.001 respectively; Fig.~\ref{props_v_lat}). 
  The best-fit models for link density and vulnerability did not include
  any effect of latitude. In contrast, the strength of the scaling relationship
  for generality increased weakly towards the poles in estuarine, marine, and
  terrestrial food webs ($\beta_{Latitude}$=0.003, $p$=0.486) but decreased 
  towards the poles in lakes and streams ($\beta_{Latitude:Lake}$=-0.019, $p$=0.027
  and $\beta_{Latitude:Stream}$=-0.031, $p$=0.002; Fig~\ref{BIT}).

  In contrast to the generally positive scaling relationships above, link
  density, generality, and vulnerability decreased as the proportion of top
  predators increased ($\beta_0$=-0.584, $p$\textless0.001; $\beta_0$=-0.418,
  $p$\textless0.001; and $\beta_0$=-0.585, $p$\textless0.001 respectively;
  Fig~\ref{props_v_lat}). For estuarine, lake, and marine food webs the
  strength of this scaling did not vary significantly with latitude
  ($\beta_{Latitude}$=0.001, $p$=0.652 for link density;
  $\beta_{Latitude}$=0.001, $p$=0.654 for vulnerability) except in stream food
  webs, where scaling was more strongly negative in the tropics and near zero
  near the poles ($\beta_{Latitude:Stream}$=0.010, $p$=0.006 for link density;
  $\beta_{Latitude:Stream}$=0.010, $p$=0.007 for vulnerability;
  Fig~\ref{BIT}). The best-fit model for generality did not include any terms
  for latitude in any ecosystem type). In stream food webs, scaling was more
  strongly negative in the tropics and did not differ from zero near the poles

@@

  The lack of a latitudinal gradient in species richness in this dataset
  contrasts strongly with other
  studies~\citep{Schemske2009,Macpherson2002,Kaufman1995}. As numbers of
  species and links included in a food web vary strongly with sampling effort
  as well as with the underlying diversity of the study area, it is possible
  that the lack of latitudinal trends here is a result of researchers tending
  to expend similar amounts of sampling effort across studies. This could
  result in food webs describing species-rich tropical communities omitting
  more species and links than species-poor arctic communities. In addition, it
  is worth noting that gradients in species richness are generally measured
  for a single taxonomic group at a time
  (e.g.~\citep{Schemske2009,Macpherson2002,Kaufman1995}). It is possible that
  these taxa are not well-represented in our food webs and that the dominant
  taxa in them do not have an underlying latitudinal gradient in richness.  In
  either case, the lack of a strong association between species richness and
  latitude in any ecosystem type means that any effect of latitude on other
  scaling  relationships is not being driven by an underlying  latitudinal
  gradient in species richness. The lack of confounding effects of latitude
  allows us to  more clearly assess effects of latitude on scaling with
  species richness and proportions of species in different  trophic levels.
